Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
updated josie slides

---
title: "CMSAC_reveal_js"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
---

@@

::::: columns
::: {.column width="50%"}
```{r}
knitr::include_graphics("4 bracket good.png")
```
:::

::: {.column width="50%"}
::: {.fragment}
```{r}
knitr::include_graphics("Mens-Basketball-Olympics-Getty-H-2024.jpeg")
```

:::

::: {.fragment}

```{r}
knitr::include_graphics("five_rec_league_men_playing_basketball.png")
```

:::

::: {.fragment}

```{r}
knitr::include_graphics("jv_girls_basketball.png")
```

:::

::: {.fragment}

```{r}
knitr::include_graphics("five_babies_playing_basketball.png")

```
:::

:::
:::::

@@

## Introduction

<!-- Overall comments 
I think the presentation is too long for 12 minutes (+ questions), but
I feel like I didn't cover everything we want. What is plausable to cover
in a 12-minute period? 
Also, we need to figure out the notation for the final metric because I think
the weight should have the subscript m, not I. And I probably was not consistent
with the usage of theta or calling it true rank/true strength. -->

-   There exists numerous competition formats across all professional sports
-   Which competition structure is best?
-   The "best" competition structure depends on how well it reflects the competitors' true strengths
-   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors' true strengths

<!-- Should we talk about Efficacy vs Effectivity here -->

## Extreme Tournament

-   4 team single elimination tournament:
    -   Dream Team vs Work Group vs Middle School Girls Team vs Babies
-   If we only care about the top team winning, seeding structure does not matter
-   If we care about who gets second and third, seeding becomes important
    -   Would not pair up Work Group and Dream Team in the first round

@@

-   Other competitions where the correct ranking of some, but not all, of the competitors is of import
    -   Olympic competition ranks the top 3 competitors
-   "Salient ranks": the ranks of a competition structure that matter
    -   Efficacy: 1 through $n$ matter
    -   Effectivity: 1 is only salient rank
-   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

>>>>

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
---

@@

::::::::: columns

::::::: {.column width="40%"}
::: {.fragment}
```{r}
knitr::include_graphics("Mens-Basketball-Olympics-Getty-H-2024.jpeg")
```
:::

::: {.fragment}
```{r}
knitr::include_graphics("five_rec_league_men_playing_basketball.png")
```
:::

::: {.fragment}
```{r}
knitr::include_graphics("jv_girls_basketball.png")
```
:::

::: {.fragment}
```{r}
knitr::include_graphics("five_babies_playing_basketball.png")

```
:::
:::::::

::::::: {.column width="60%"}
::: {.fragment}
```{r, out.width="80%"}
knitr::include_graphics("4 bracket good.png")
```
:::
:::::::
:::::::::

@@

::: notes
-   4 team single elimination tournament:
    -   Dream Team vs Work Group vs Middle School Girls Team vs Babies
-   If we only care about the top team winning, seeding structure does not matter
-   If we care about who gets second and third, seeding becomes important
    -   Would not pair up Work Group and Dream Team in the first round
:::

@@

-   Decide how many rankings are important
    -   Olympic competition ranks the top 3 competitors
    -   Premier League: top 4 (Champions League) and bottom 3 (Relegation)
-   **Salient ranks**: the ranks that matter
    -   Efficacy: 1 through $n$ matter
    -   Effectivity: 1 is only salient rank

## Goal

-   Want to quantify how well a structure can accurately order the competitors
-   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

## Distribution of $\hat{r}$

```{r}
knitr::include_graphics("r_dist.png")
```
